### 2004-04-28 Documentation/part2.tex
Documentation part2.tex

\part{The Experiment Definition Module (\EDM )}

\chapter{Basics: Source, Oscillation, Detection}

Here comes how a experiment is defined (what is a channel, a rule, how are 
they treated, ...). Splitting in source, oscillation, and detector, etc.

\chapter{Getting started: Editing existing experiments}

Here comes how to load and modify existing experiments;
also: how does the main screen work and the most important functions at 
exampled

\chapter{Features of the \EDM }

\chapter{General concept}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Source properties}
\label{sec:source}

@@

In the latter case the fluxes are read from a file with seven rows and
501 lines.
\begin{quotation}
$ E\quad
\Phi_{\nu_e}\quad
\Phi_{\nu_\mu}\quad
\Phi_{\nu_\tau}\quad
\Phi_{\bar\nu_e}\quad
\Phi_{\bar\nu_\mu}\quad
\Phi_{\bar\nu_\tau}$
\end{quotation}
where the energy values are evenly spaced. 
For accessing fluxes at 
arbitrary energies linear interpolation is used. Once the energy leaves the
range of values given in the file $0.0$ is returned. The units used for the
fluxes are not specified and right now each flux has its own fudge factor
which is hard encoded in the file {\tt nufact.c}. The problem is that
in computing the normalization one has to know things like number of target
particles per unit mass asf. which is not always straitforward. Thus a fudge
factor will be usefule in any case. Right now however this is not the only 
fudge factor since for some experiments the definition of their target mass
involves another fudge factor.

@@

For the parameterization of an energy calibration error two possibilities
are implemented. The first one being somewhat simpler, whereas the second one
is more accurate but requires a careful choice of parameters. The first 
option is
\begin{equation}
s_i(a,b):=s_i(a)+b\cdot s_i\, E_i/(E_\mathrm{max}-E_\mathrm{min})
\end{equation}
and it is refered to as tilt since it describes a linear distortion, a tilt
of the event rate spectrum. The second option is closer to an actual energy
calibration error, which basicaly amounts to replacing the reconstructed 
energy $E'$ by $(1+b)\cdot E'$ and we use following approximation
\begin{eqnarray}
s_i(b)&=& (1+b)\cdot(s_{k+1}-s_k)\cdot\delta+s_i\,,\\
\delta&=&b\cdot(i+\Delta E+ 1/2)+i\,,\nonumber\\
k&=&\mod(\delta,1)\,.\nonumber
\end{eqnarray}
Special care is required at the boundaries when $k<1$ or $k+1>N_\mathrm{bins}$,
since there $s_k$ or $s_{k+1}$ may not have been calculated. One possibility
to deal with those cases is to assume that for those values of $k$ $s_k$ is
zero. This is the default. In cases where $s$ at the boundary is however still
sizeable this may lead to untolerable errors and subsequently to a wrong
estimate of the impact of a calibration error. Therefore in those cases it
is advisable to truncate the analysis range by a few bins at the boundary
and therefore ensure in this way that only those $s_i$ are used whose index
$k$ is within the range $1,\ldots, N_\mathrm{bins}-1$.

@@

Before going into the details we define that the range from $E_\mathrm{min}$
to $E_\mathrm{max}$ is diveded in a certain number of bins $N_\mathrm{bins}$.
These bins form a partition, which not necessarily is equidistant. The 
partition is any case completely defined by specifying $E_\mathrm{min}$
and $E_\mathrm{max}$ and the size of each bin  $\Delta E_i$. It is also
straightforward to compute the midpoint of each bin $E_i$.
In order to obtain the number of events 
$n_i^c$ from one channel $c$ in one bin $i$ one has to solve this integral
%
\begin{eqnarray}
\label{eq:events_bin}
n_i^c=N/L^2\,\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
\quad \int dE \,\, \Phi^c(E)\,
P^c(E)\,
\sigma^c(E)\,
R^c(E,E')\,
\epsilon^c(E')\,.
\end{eqnarray} 
%
The integration with respect to the reconstructed energy $E'$ can be
performed independently of the oscillation parameters and we define
the bin kernel for the $i$th bin $K_i^c$
\begin{equation}
\label{eq:kernel}
K_i^c(E):=\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
\quad R^c(E,E')\,
\epsilon^c(E')\,.
\end{equation}
With this definition equation~\ref{eq:events_bin} can be rewritten as
\begin{equation}
\label{eq:simple_int}
n_i^c=N/L^2 \int dE\quad  \underbrace{\Phi^c(E)\,
P^c(E)\,
\sigma^c(E)\,
K_i^c(E)\,}_{f(E)}. 
\end{equation}

@@

The range for the summation $k_l^i$ and $k_u^i$ is given by the range of 
integration $a_i$ and $b_i$ in equation~\ref{eq:simple_int}, the integration 
limits there are $0$ to $\infty$. For practical purposes however the range
can be reduced to $a_i$ and $b_i$. They are  determined by requiring
that the area under the bin kernel $K_i^c$ has reached a specific 
value close to one,
the so called confidence level $c$ (new name would be better). $1-c$ is a 
small number and specifies the fraction of  events in bin which lie outside
the summation range, usually $1-c=10^{-3}$ is sufficient. Furthemore the
value of the bin kernel at the lower and upper boundary should
be the same
\begin{eqnarray}
\int_{a_i}^{b_i} dE\quad K_i^c(E) &=& c\,,\\
K_i^c(a_i)&=&K_i^c(b_i)\,.
\end{eqnarray} 
Knowing $a_i$ and $b_i$ it is easy to obtain $k_l$ and $k_u$ by requiring
that the summation extends one bin beyond  each $a_i$ and $b_i$. Furthermore
it is possible to specify an offset which extends the index range on each side
by its value.

Since the $a_i$'s and $b_i$'s are not known in advance it could be possible
that $a_0$ and/or $b_{N_\mathrm{bins}}$ lie outside the computational domain,
\eg\ no cross section or flux is defined at those energies. To remidy this 
problem the computational domain can be specified. No calculation in \GLOBES\
will leave this domain.

In order to ensure that fast oscillating probablities do not lead to aliasing
it is desirable to impose a low-pass filter already during the calculation
of the probablities itself. This is possible and implemented has a highly
experimental feature called ``filter''\index{Filter}. 
The calculation of oscillation
probabilities basically is a computation of phase differences. Restricting
the maximum admissable size of those phase difference effectively filters
the high frequency component of the oscillation probability. This idea is
implemented according to
\begin{eqnarray}
\label{eq:filter_a}
P_{\alpha\beta}(E)&=&\sum_{ij}
U_{\alpha j} U^*_{\beta j} U^*_{\alpha i} U_{\beta i} 
e^{-i\Phi_{ij}}\times 
e^{ -\Phi_{ij}^2/\sigma_f(E)^2 }\,,
\end{eqnarray}
where $\Phi_{ij}:=\Delta m_{ij}^2 L/2E$ is the usual phase difference and
the last term is a Gau\ss ian filter with width $\sigma_f(E)$. Choosing
$\sigma_f(E):=\sigma_f^0 \cdot E$ ensures that this filter behaves 
approximately like an energy resolution function with constant width 
$\sigma_e=\sqrt{2}/\sigma_f^0$, \ie\
\begin{equation}
\label{eq:filter_b}
\int d\tilde E\quad P(\tilde E) \frac{1}{\sigma_e\,\sqrt{2\pi}}\,
e^{-\frac{(E-\tilde E)^2}{2\sigma^2_e}}\,.
\end{equation}
The equivalence of equation~\ref{eq:filter_a} and equation~\ref{eq:filter_b}
is not obvious and connected to the properties of $P_{\alpha\beta}$, 
see~\cite{Kiers:1996zj,Giunti:2003ax}. This feature works \emph{only} 
for vacuum and constant denisties and controled
by the filer state variable and $\sigma_f^0$ is set by the filter value 
variable. Assuming that $P_{\alpha\beta}$ still is the source for the highest
frequencies one can splitt the energy resolution function $\sigma(E)^c$ in
two parts by
\begin{equation}
\sigma_c^2(E)=\underbrace{\sigma_c^2(E)-\sigma_c^2(E_\mathrm{min})}_
{\tilde\sigma^2_c(E)}+\underbrace{\sigma_c^2(E_\mathrm{min})}_{\sigma_e^2}\,,
\end{equation}
where $\tilde\sigma_c(E)$ is used instead of $\sigma_c(E)$ in computing the
smearing data and $\sigma_e$ is used to determine $\sigma_f^0$.

Finally the user has the possibility to directly specify the smearing data.
The format requires to provide $k_l^i$ and $k_u^i$ for each bin, the number of
bins, the number of sampling points, the lower and upper limit of the sampling
range. Furthemore the numbers $K_i(E_j)^c$ have to provide for all values
in between $k_l^i$ and $k_u^i$. The convolution is then performed according to
equation~\ref{eq:algo_one}.

\chapter{Creating new experiments}

Most advanced step: How to create a new experiment from scratch .

>>>>

\part{The Abstract Experiment Definition Language -- AEDL}

\chapter{General Concept}

The aim of AEDL is to describe a large variety of complex experiments
by a limited number of parameters and to allow a representation of
very different setups within one data structure, thus enabling a
universal rate and $\chi^2$ computation. Usually in dealing with
experiment simulations a new piece of code is written and compiled
for each different experiments. In many cases even changes of some
parameter like the number of bins requires the recompilation of
large portions of the source code. This technique may however soon
reach its limits in cases where the simulated experiments are rather complex
and/or more than one kind of experiment is going to examined. Furthermore
is is very difficult to assure the correctness of the results obtained
in this way since every time a new piece of code is added to handle a new
experiment new errors will be introduced.

Thus a general and flexible experiment description language is needed. 
A neutrino experiment can be split into three parts: source, oscillation
and detection. The sources within \GLOBES\ are assumed to be stationary
point sources and each experiment has only one source. Specifically this
does not allow to study cases where the time dependence of the neutrino flux
is of physical significance like \eg\ in supernov\ae. It however is possible
to study beams which have a bunch structure, since the time dependence of the
neutrino source is physically only in as far important as it is used to
suppress backgrounds. Also geometrical effects of a source distribution like
\eg\ in the Sun can not described in this way. For this kind of sources
the oscillation part of an 
experiment is simply
given by the baseline and the matter density profile.

While
it is comparatively simple to define a general neutrino source 
and the oscillation physics, the general properties of a detector are more
complicated. The basic assumption in building an abstract detector description
is \emph{linearity}, \ie\ two neutrino events do not interfere with each other.
Furthermore it is assumed that all information on the oscillation physics 
is given by the \emph{reconstructed} flavour and energy of a 
neutrino event. The term ``reconstructed'' implies that there is not a single
well defined value for each event and variable but a distribution of possible
values. This illustrated in figure~\ref{fig:distro} for the energy variable.
The same is of course possible with the flavour variable. In that case however
only discrete values are possible.
%
\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.6\textwidth]{mapping}
\end{center}
\caption{\label{fig:distro} A detector maps the true values into
a distribution of reconstructed values of a parameter. Here the energy is 
shown.}
\end{figure}
%

This picture also can be formulated in a more mathematical language. Let $x$
be the true variable and $x'$ the reconstructed variable then $f(x)$ is the
distribution of true values and $p(x')$ is the distribution of reconstructed
values. Then the detector function  $D(x,x')$ which describes 
the mapping performed by the detector is defined by
\begin{eqnarray}
\label{eq:mapping}
p(x')&=&\int dx\, f(x)\cdot D(x,x')\,.
\end{eqnarray}
Obviously equation~\ref{eq:mapping} can only describe the detector correctly
when the condition of linearity is fulfilled. Within this model a detector
is completely specified by a set of $D(E,E')$ for the energy variable $E$
and a set $D(F,F')$ for the flavour variable $F$, where usually $D(E,E')$ also
depends on the true flavour $F$ as well as $D(F,F')$ depends on the true energy
$E$. Those sets of mapping functions usually will be obtained by a 
full detector simulation and can be obtained by using as input 
distribution $f(x)$ a delta distribution $\delta(x-x_0)$.

In order to allow a systematic experiment definition including various
sources of systematical errors we use several layers. The first layer is
the so called ``channel'', it is the link between the oscillation physics
and the detector. A channel specifies how a certain neutrino type in the 
source is mapped into a reconstructed neutrino type, \eg\ a muon neutrino
oscillates into an electron neutrino and subsequently interacts via
quasi-elastic charged current scattering. The measured energy and direction of
the electron in the detector finally allow to reconstruct the neutrino energy.
The channels are the building blocks for so called ``rules'' which
contain information like the efficiencies and systematical errors. An example
could look like this: the efficiency for quasi-elastics electron neutrino
events is $0.4$. There is a fraction of $0.01$ of all neutral current events
which are mis-identified as quasi-elastic electron neutrino events. 
The neutral current fraction is only known within $10\%$ and there is
an energy scale uncertainty of $100\,\mathrm{MeV}$. Thus a rule connects the
event rates to the calculation of a $\chi^2$ which properly includes 
systematical errors. This $\chi^2$ is then the starting point for the
oscillation physics analysis.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Getting started}

As a first example we discuss how to define a neutrino factory experiment using
AEDL. The first line of every experiment definition file is
\begin{quote}
{\tt !\%GLoBES}
\end{quote}

We tell \GLOBES\ that we want to use the built-in neutrino factory
flux, by setting the beam type variable to $1$. The target mass 
is $50\,\mathrm{kt}$ and a total of $ 4.264\cdot10^{21}$
muons is stored. The data is divided into 20 energy bins ranging from
$4\,\mathrm{GeV}$ to $50\,\mathrm{GeV}$. 
\begin{quote}
{\tt /* beam */}\\
{\tt \$beam\_type = 1}\\
{\tt \$target\_mass = 50}\\
{\tt \$stored\_muons = 4.264e+21}\\
{\tt \$numofbins = 20}\\
{\tt \$emin = 4.0}\\
{\tt \$emax = 50.0}
\end{quote}

Next a file has to be specified which contains the cross section we want to 
use.
\begin{quote}
{\tt /* cross section */}\\
{\tt cross(\#CC)<}\\
{\tt \hspace*{3ex}@cross\_file = XCC.dat}\\
{\tt >}
\end{quote}
The command {\tt cross} tells the parser that a cross section environment
begins and has the name {\tt \#CC}. These names are used to refer later on
to this specific environment and each name begins with a leading {\tt \#}.
Whenever we need to specify a cross section from here on we can use {\tt \#CC}
to refer to the cross section as given in the file {\tt XCC.dat}.

Of course also the baseline and matter profile have to be specified, which
 looks like this
\begin{quote}
{\tt /* baseline */}\\
{\tt \$baseline = 3000.0}\\
{\tt \$densitytab = \{3.5\}}\\
{\tt \$lengthtab = \{3000.0\}}\\
{\tt \$density\_error = 0.05}
\end{quote}
The curly brackets used for the definition of {\tt \$densitytab} and
{\tt \$lengthtab} delimit as simple list of numbers -- here the lists
have length $1$. {\tt \$density\_error} is the relative $1\,\sigma$ uncertainty
on the density.

The energy resolution function can easily be specified with
\begin{quote}
{\tt /* energy resolution */}\\
{\tt energy(\#MINOS)<}\\
{\tt \hspace*{3ex}@algorithm = 1}\\
{\tt \hspace*{3ex}@bins = 20}\\
{\tt \hspace*{3ex}@emin = 4.0}\\
{\tt \hspace*{3ex}@emax = 50.0}\\
{\tt \hspace*{3ex}@energy\_resolution = \{0.15,0.0,0.0\}}\\
{\tt >}
\end{quote}
The {\tt energy} command opens the corresponding energy environment and
the energy resolution function defined here is refered to by the name 
{\tt \#MINOS}. Out of several possibilities it employs algorithm $1$,
the simplest and fastest one. The actual energy resolution is specified
by the energy resolution variable which is a list of three elements. Each 
element is one parameter of a general resolution function as defined in 
equation ~\ref{eq:sigma_e}.

Now we define the one appearance and the corresponding disappearance 
channel of a neutrino factory: 
$\nu_e\rightarrow\nu_\mu$  and $\bar\nu_\mu\rightarrow\bar\nu_\mu$ 
($\mu^+$ stored).
\begin{quote}
{\tt /* channels */}\\
{\tt channel(\#appearance)<}\\
{\tt \hspace*{3ex}@channel = +: +: electron: muon: \#CC: \#MINOS}\\
{\tt >}\\
{\tt channel(\#disappearance)<}\\
{\tt \hspace*{3ex}@channel = +: -: muon: muon: \#CC: \#MINOS}\\
{\tt >}
\end{quote}
The first $\pm$ selects the polarity, \ie\ wether $\mu^+$ or $\mu^-$ are 
stored. The second $\pm$ determines whether neutrinos or anti-neutrinos are
taken from the flux table. The third position defines the initial flavour
and the forth position the final flavour, followed by the name of the cross
section and energy resolution function.

The last step is to assemble the channels into a rule.
\begin{quote}
{\tt /* rules */}\\
{\tt rule(\#rule1)<}\\
{\tt \hspace*{3ex}@signal = 0.45 @ \#appearance}\\
{\tt \hspace*{3ex}@signalerror = 0.001 : 0.0001}\\
{\tt \hspace*{3ex}@background = 1.0e-05 @ \#disappearance}\\
{\tt \hspace*{3ex}@backgroundcenter = 1 : 0.0}\\
{\tt \hspace*{3ex}@backgrounderror = 0.05 : 0.0001}\\
{\tt \hspace*{3ex}@errordim = 0}\\
{\tt \hspace*{3ex}@energy\_window = 4.0 : 50.0}\\
{\tt >}
\end{quote}
The signal variable says that the signal in our experiment is the 
channel named {\tt \#appearance} with an constant efficiency of $0.45$.
The signal error variable has two components: the first one is the
normalization error of the signal, here it is $0.1\%$. The second component
signifies the energy calibration error of the signal, whose definition 
can be found in section~\ref{sec:energy}. The background variable
specifies the composition of the beam background and this case is
given by the channel named {\tt \#disappearance}, \ie\ this are
the muons whose charge has been mis-identified. They enter the calculation
with an efficiency of $1\cdot 10^{-5}$. The background center variable allows
to rescale the total background contribution which is especially useful if
there are many background components. 
The background error variable is defined in the same way as the signal error
variable, \ie\ we have a $5\%$ background uncertainty.
The error dimension variable selects how
the systematical errors are treated according to table~\ref{tab:error_dim}.

The above defined experiment represents a first version of a neutrino factory
experiment. It still lacks the correct energy dependence of the efficiencies
and of course the channels and rules for the operation with $\mu^-$ stored

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Features of the  AEDL}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Source properties}
\label{sec:source}

@@

In the latter case the fluxes are read from a file with seven rows and
501 lines.
\begin{quotation}
$ E\quad
\Phi_{\nu_e}\quad
\Phi_{\nu_\mu}\quad
\Phi_{\nu_\tau}\quad
\Phi_{\bar\nu_e}\quad
\Phi_{\bar\nu_\mu}\quad
\Phi_{\bar\nu_\tau}$
\end{quotation}
where the energy values are evenly spaced. 
For accessing fluxes at 
arbitrary energies linear interpolation is used. Once the energy leaves the
range of values given in the file $0.0$ is returned. The units used for the
fluxes are not specified and right now each flux has its own fudge factor
which is hard encoded in the file {\tt nufact.c}. The problem is that
in computing the normalization one has to know things like number of target
particles per unit mass asf. which is not always straightforward. Thus a fudge
factor will be useful in any case. Right now however this is not the only 
fudge factor since for some experiments the definition of their target mass
involves another fudge factor.

@@

For the parameterization of an energy calibration error two possibilities
are implemented. The first one being somewhat simpler, whereas the second one
is more accurate but requires a careful choice of parameters. The first 
option is
\begin{equation}
s_i(a,b):=s_i(a)+b\cdot s_i\, E_i/(E_\mathrm{max}-E_\mathrm{min})
\end{equation}
and it is refered to as tilt since it describes a linear distortion, a tilt
of the event rate spectrum. The second option is closer to an actual energy
calibration error, which basically amounts to replacing the reconstructed 
energy $E'$ by $(1+b)\cdot E'$ and we use following approximation
\begin{eqnarray}
s_i(b)&=& (1+b)\cdot(s_{k+1}-s_k)\cdot\delta+s_i\,,\\
\delta&=&b\cdot(i+\Delta E+ 1/2)+i\,,\nonumber\\
k&=&\mod(\delta,1)\,.\nonumber
\end{eqnarray}
Special care is required at the boundaries when $k<1$ or $k+1>N_\mathrm{bins}$,
since there $s_k$ or $s_{k+1}$ may not have been calculated. One possibility
to deal with those cases is to assume that for those values of $k$ $s_k$ is
zero. This is the default. In cases where $s$ at the boundary is however still
sizeable this may lead to intolerable errors and subsequently to a wrong
estimate of the impact of a calibration error. Therefore in those cases it
is advisable to truncate the analysis range by a few bins at the boundary
and therefore ensure in this way that only those $s_i$ are used whose index
$k$ is within the range $1,\ldots, N_\mathrm{bins}-1$.

@@

Before going into the details we define that the range from $E_\mathrm{min}$
to $E_\mathrm{max}$ is divided in a certain number of bins $N_\mathrm{bins}$.
These bins form a partition, which not necessarily is equidistant. The 
partition is any case completely defined by specifying $E_\mathrm{min}$
and $E_\mathrm{max}$ and the size of each bin  $\Delta E_i$. It is also
straightforward to compute the midpoint of each bin $E_i$.
In order to obtain the number of events 
$n_i^c$ from one channel $c$ in one bin $i$ one has to solve this integral
%
\begin{eqnarray}
\label{eq:events_bin}
n_i^c=N/L^2\,\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
\quad \int dE \,\, \Phi^c(E)\,
P^c(E)\,
\sigma^c(E)\,
R^c(E,E')\,
\epsilon^c(E')\,.
\end{eqnarray} 
%
The integration with respect to the reconstructed energy $E'$ can be
performed independently of the oscillation parameters and we define
the bin kernel for the $i$th bin $K_i^c$
\begin{equation}
\label{eq:kernel}
K_i^c(E):=\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
\quad R^c(E,E')\,
\epsilon^c(E')\,.
\end{equation}
With this definition equation~\ref{eq:events_bin} can be rewritten as
\begin{equation}
\label{eq:simple_int}
n_i^c=N/L^2 \int dE\quad  \underbrace{\Phi^c(E)\,
P^c(E)\,
\sigma^c(E)\,
K_i^c(E)\,}_{f(E)}. 
\end{equation}

@@

The range for the summation $k_l^i$ and $k_u^i$ is given by the range of 
integration $a_i$ and $b_i$ in equation~\ref{eq:simple_int}, the integration 
limits there are $0$ to $\infty$. For practical purposes however the range
can be reduced to $a_i$ and $b_i$. They are  determined by requiring
that the area under the bin kernel $K_i^c$ has reached a specific 
value close to one,
the so called confidence level $c$ (new name would be better). $1-c$ is a 
small number and specifies the fraction of  events in bin which lie outside
the summation range, usually $1-c=10^{-3}$ is sufficient. Furthermore the
value of the bin kernel at the lower and upper boundary should
be the same
\begin{eqnarray}
\int_{a_i}^{b_i} dE\quad K_i^c(E) &=& c\,,\\
K_i^c(a_i)&=&K_i^c(b_i)\,.
\end{eqnarray} 
Knowing $a_i$ and $b_i$ it is easy to obtain $k_l$ and $k_u$ by requiring
that the summation extends one bin beyond  each $a_i$ and $b_i$. Furthermore
it is possible to specify an offset which extends the index range on each side
by its value.

Since the $a_i$'s and $b_i$'s are not known in advance it could be possible
that $a_0$ and/or $b_{N_\mathrm{bins}}$ lie outside the computational domain,
\eg\ no cross section or flux is defined at those energies. To remedy this 
problem the computational domain can be specified. No calculation in \GLOBES\
will leave this domain.

In order to ensure that fast oscillating probabilities do not lead to aliasing
it is desirable to impose a low-pass filter already during the calculation
of the probabilities itself. This is possible and implemented has a highly
experimental feature called ``filter''\index{Filter}. 
The calculation of oscillation
probabilities basically is a computation of phase differences. Restricting
the maximum admissible size of those phase difference effectively filters
the high frequency component of the oscillation probability. This idea is
implemented according to
\begin{eqnarray}
\label{eq:filter_a}
P_{\alpha\beta}(E)&=&\sum_{ij}
U_{\alpha j} U^*_{\beta j} U^*_{\alpha i} U_{\beta i} 
e^{-i\Phi_{ij}}\times 
e^{ -\Phi_{ij}^2/\sigma_f(E)^2 }\,,
\end{eqnarray}
where $\Phi_{ij}:=\Delta m_{ij}^2 L/2E$ is the usual phase difference and
the last term is a Gau\ss ian filter with width $\sigma_f(E)$. Choosing
$\sigma_f(E):=\sigma_f^0 \cdot E$ ensures that this filter behaves 
approximately like an energy resolution function with constant width 
$\sigma_e=\sqrt{2}/\sigma_f^0$, \ie\
\begin{equation}
\label{eq:filter_b}
\int d\tilde E\quad P(\tilde E) \frac{1}{\sigma_e\,\sqrt{2\pi}}\,
e^{-\frac{(E-\tilde E)^2}{2\sigma^2_e}}\,.
\end{equation}
The equivalence of equation~\ref{eq:filter_a} and equation~\ref{eq:filter_b}
is not obvious and connected to the properties of $P_{\alpha\beta}$, 
see~\cite{Kiers:1996zj,Giunti:2003ax}. This feature works \emph{only} 
for vacuum and constant densities and controlled
by the filer state variable and $\sigma_f^0$ is set by the filter value 
variable. Assuming that $P_{\alpha\beta}$ still is the source for the highest
frequencies one can split the energy resolution function $\sigma(E)^c$ in
two parts by
\begin{equation}
\sigma_c^2(E)=\underbrace{\sigma_c^2(E)-\sigma_c^2(E_\mathrm{min})}_
{\tilde\sigma^2_c(E)}+\underbrace{\sigma_c^2(E_\mathrm{min})}_{\sigma_e^2}\,,
\end{equation}
where $\tilde\sigma_c(E)$ is used instead of $\sigma_c(E)$ in computing the
smearing data and $\sigma_e$ is used to determine $\sigma_f^0$.

Finally the user has the possibility to directly specify the smearing data.
The format requires to provide $k_l^i$ and $k_u^i$ for each bin, the number of
bins, the number of sampling points, the lower and upper limit of the sampling
range. Furthermore the numbers $K_i(E_j)^c$ have to provide for all values
in between $k_l^i$ and $k_u^i$. The convolution is then performed according to
equation~\ref{eq:algo_one}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{User-defined systematics}
\label{sec:ud_error_dim}.
